### 2025-11-19 cluster-analysis.qmd
20251119 - tidyLPA for latent profile analysis

```{r}
library("petersenlab")
library("nflreadr")
library("mclust")
library("plotly")
library("tidyverse")
```

@@

@@

@@

>>>>

```{r}
library("petersenlab")
library("nflreadr")
library("mclust")
library("tidyLPA")
library("plotly")
library("tidyverse")
```

@@

We can also use the `tidyLPA` package [@R-tidyLPA; @Rosenberg2018_packages], which provides an interface to the `mclust` package [@R-mclust; @Scrucca2023_packages].

```{r}
# model 1 (EEI): Equal variances and covariances fixed to 0
# model 2 (VVI): Varying variances and covariances fixed to 0
# model 3 (EEE): Equal variances and equal covariances
# model 4      : Varying variances and equal covariances (not able to be fit w/ mclust but can be fit using package = "mplus", if installed)
# model 5      : Equal variances and varying covariances (not able to be fit w/ mclust but can be fit using package = "mplus", if installed)
# model 6 (VVV): Varying variances and varying covariances

wrTypes_lpa_classes <- tidyLPA::estimate_profiles(
  df = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
  n_profiles = 1:6,
  models = c(1, 2, 3) #c(1, 2, 3, 6) takes too long to run because of the model with varying variances and varying covariances (model 6)
)

wrTypes_lpa_classes

tidyLPA::compare_solutions(
  wrTypes_lpa_classes,
  statistics = c("AIC", "BIC", "Entropy", "BLRT_p"))

wrTypes_lpa <- tidyLPA::estimate_profiles(
  df = merged_dataWRs_cluster %>% select(-gsis_id, -display_name),
  n_profiles = numTypesWR,
  model = 3 # equal variances and equal covariances
)

wrTypes_lpa
tidyLPA::get_fit(wrTypes_lpa)
```

@@

```{r}
tidyLPA::plot_profiles(wrTypes_lpa)
```

@@

And here are the clusters based on the `tidyLPA` [@R-tidyLPA; @Rosenberg2018_packages] solution:

```{r}
merged_dataWRs_cluster <- cbind(
  merged_dataWRs_cluster,
  tidyLPA::get_data(wrTypes_lpa) %>% 
    select(model_number, classes_number, CPROB1:Class)
)

merged_dataWRs_cluster %>% 
  select(display_name, CPROB1:Class)

merged_dataWRs_cluster %>% 
  group_by(Class) %>% 
  summarise(across(
    where(is.numeric),
    ~ mean(., na.rm = TRUE)
    )) %>% 
  t() %>% 
  round(., 2)
```

### 2025-11-19 getting-started.qmd
20251119 - tidyLPA for latent profile analysis

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","patchwork","pROC",
     "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust","magrittr",
     "viridis","viridisLite","msir","plotly","webshot2","quantmod","fPortfolio",
     "NMOF","nFactors","xts","zoo","forecast","parallelly","doParallel",
     "missRanger","ggridges","powerjoin","bestNormalize","LongituRF","gpboost",
     "mgcv","rms","car","lavaan","lavaanPlot","lavaangui","mice","miceadds",
     "interactions","robustbase","ordinal","MASS","data.table","future",
     "future.apply","SimDesign","domir","GGally","Rglpk","TTR"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

>>>>

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","patchwork","pROC",
     "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust","magrittr",
     "viridis","viridisLite","msir","plotly","webshot2","quantmod","fPortfolio",
     "NMOF","nFactors","xts","zoo","forecast","parallelly","doParallel",
     "missRanger","ggridges","powerjoin","bestNormalize","LongituRF","gpboost",
     "mgcv","rms","car","lavaan","lavaanPlot","lavaangui","mice","miceadds",
     "interactions","robustbase","ordinal","MASS","data.table","future",
     "future.apply","SimDesign","domir","GGally","Rglpk","TTR","rockchalk",
     "yhat","tidyLPA"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

### 2025-11-19 index.qmd
20251119 - tidyLPA for latent profile analysis

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
  "domir", "GGally","Rglpk","TTR",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic", "r4ds.tutorials" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "bestNormalize", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", 
  "lavaan", "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", 
  "ordinal", "MASS", "data.table", "future", "future.apply", "SimDesign", 
  "domir", "GGally","Rglpk","TTR","rockchalk","yhat","tidyLPA",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic", "r4ds.tutorials" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)
